---
title: "Lab 7: Team All Stars"
author: "Johnny Tamanaha, Emily Archambault, Eugene Miller, Kaiyuan Fan, Jingzhi Zhang"
date: "10/10/2019"
output: html_document
---

### Dataset Description
This dataset contains 2075259 measurements gathered in a house located in Sceaux (7km of Paris, France) between December 2006 and November 2010 (47 months). It includes seven different power measurements as well as date and time. We handled missing values by removing them from the dataset, as we could not justify their significance. 

```{r, setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
power_data <- read_delim("household_power_consumption.txt", delim = ";", na = c('', '?'),
                  col_types = cols(Date = col_date("%d/%m/%Y"),
                                   Global_active_power = col_double(),
                                   Global_reactive_power = col_double(),
                                   Global_intensity = col_double(),
                                   Sub_metering_1 = col_double(),
                                   Sub_metering_2 = col_double(),
                                   Sub_metering_3 = col_double()),
                                    locale = locale(decimal_mark = ".", grouping_mark = ","))
```

## Individual Sections

## Johnny Tamanaha
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
data <- read_delim("household_power_consumption.txt", delim=";", na = c("","?"), col_types=cols(Date = col_date("%e/%m/%Y"), Time=col_time("%H:%M:%S"), .default=col_double()), locale=locale(decimal_mark="."))
```
```{r}
data_year <- data %>%
  separate(Date, into=c("year", "month", "day"), "-")
avg_global_intensity <- data_year %>%
  group_by(year, Time) %>%
  summarize(avg = mean(Global_intensity, na.rm=TRUE)) %>%
  filter(!is.na(year))
ggplot(avg_global_intensity) +
  geom_point(mapping=aes(x=Time, y=avg, color=year)) +
  labs(title="Average Household Active Power Split by Year") +
  ylab("Active Power (amp)") +
  xlab("Time (Hour:Minutes:Seconds)") +
  scale_color_discrete(name=("Year"))
```
```{r}
avg_global_intensity_stat <- data_year %>%
  group_by(year) %>%
  summarize(avg = mean(Global_intensity, na.rm=TRUE)) %>%
  filter(!is.na(year))
avg_global_intensity_stat
```

Question: How has daily power usage changed from 2006 to 2010?

I first separated the date variable into year, month and day. Then, I grouped the data by year and time and summarized to find the average power usage at each time of each year. Finally, I plotted the time vs the average power and split it by year. To find a summary statistic, I grouped by year and took the average power usage over each year.

My question is interesting because it shows how the general power usage for the people of France changes over time. 

At first I was trying to compare different variables with time between 2006 and 2010. The graphs always looked too cluttered to gain any insight, so I decided to average the variables I chose at each point in time for each year, so that the points would be smoother to look at. After I found the variable that I wanted, I realized that there weren't too many years to plot in the entire data set, so I decided to stick them all on the same plot to show a better time series of the power usage. 

I think that renewable energy companies benefit from my analysis because it shows a slight trend towards less power usage. The exact opposite goes for power companies. 

## Eugene Miller

There are 11 official holidays in France, where this dataset was created. How does energy use vary depending on whether or not it is a holiday? Three holidays are on 8 May, 25 December, 14 July (V-Day, Christmas, and Bastille Day). By comparing these holidays to control days five days prior to the holiday, we can determine whether these holidays constitute abnormal energy usage. 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
data <- power_data %>%
  separate(Date,into = c("Year", "Month", "Day"), convert = TRUE) %>%
  separate(Time, into = c("Hour", "Minute", "Second"), convert = TRUE) %>%
  mutate(Active_energy_consumed = (Global_active_power*1000/60 - Sub_metering_1 - Sub_metering_2 - Sub_metering_3), Datetime = make_datetime(Year, Month, Day, Hour, Minute, Second)) 
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}

holidays1 <- data %>%
  filter(`Day` == 8 & `Month` == 4) 
holidays2 <- data %>%
  filter(Day == 25 & Month == 12)
holidays3 <- data %>%
  filter(Day == 14 & Month == 7)
control1 <- data %>%
  filter(Day == 3 & Month == 4)
control2 <- data %>%
  filter(Day == 20 & Month == 12 & !is.na(Active_energy_consumed))
  
control3 <- data %>%
  filter(Day == 9 & Month ==7)

holiday_data <- rbind(holidays1,holidays2,holidays3,control1,control2,control3)

#These charts show both higher and more sustained energy usage during holidays as comared to a normal day.
ggplot(holiday_data, aes(Hour, Active_energy_consumed)) +
  geom_col() + 
  facet_wrap(Month~Day) +
  labs(x = "Hour of Day", y = "Energy Consumed (Watt Hours)")

hsum1 <- sum(holidays1[14]) / 1000
csum1 <- sum(control1[14]) / 1000
hsum2 <- sum(holidays2[14]) / 1000
csum2 <- sum(control2[14]) / 1000
hsum3 <- sum(holidays3[14]) / 1000
csum3 <- sum(control3[14]) / 1000

```

These charts show both higher and more sustained energy usage during holidays as comared to a normal day. We can demonstrate this difference statistically as well. Comparing April 3: `r csum1` kWh to April 8: `r hsum1` kWh, December 20: `r csum2` kWh to December 25: `r hsum2` kWh, and July 9: `r csum3` kWh to July 14: `r hsum3` kWh we see the higher use on holidays. 

This is an interesting question because it shows how factors outside the dataset can affect measurements that we percieve in the dataset. To come to this question, I first plotted energy use over month, split by night and day. Noticing variation over the year, I decided it would be interesting to look at specific holidays and see how their energy use varied compared to normal days.

I think this analysis would be useful to climate change activists who want to show that we are using too much energy, and it would be harmful to power companies who try to downplay their influence in people's lives.

##Emily Archambault

Question: How does power usage from laundry room devices vary across the months of the year?

``{r, echo = FALSE, message = FALSE, warning = FALSE}
washmonth <- power_data %>% select(month, Sub_metering_2) %>% group_by(month) %>% 
summarise(mean.sub2=mean(Sub_metering_2, na.rm=TRUE))
ggplot(washmonth) + geom_col(aes(month, mean.sub2)) + 
labs(x = "Month", y = "Sub Metering 2 (Watt-Hours)", title ="Laundry Room Power Usage")
```
```{r}
janmean <- data1 %>% filter(month == 1) %>% select(Sub_metering_2)%>% 
summarise(mean.jan=mean(Sub_metering_2, na.rm=TRUE))
augmean <- data1 %>% filter(month == 8) %>% select(Sub_metering_2)%>% 
summarise(mean.aug=mean(Sub_metering_2, na.rm=TRUE))
```
From looking at this graph, it appears that average power use (in watt-hours) for the laundry room is higher during the winter months, and drops lower in the summer. I plotted the mean value for Sub Metering 2 across each month using a column chart. This is supported by calculating the exact value of the mean for different months and comparing. For example, in January it is `r janmean`, and in August it is `r augmean`.

This question is interesting because it shows a pattern in the use of laundry room power, and that it is not simply constant across a year like one might initially think. I came up with this question by looking at the different sub metering categories, and thinking of reasons the power usage for each one might change over time, and then plotting some graphs to see if any of my ideas were correct. 

Laundry detergent and other various laundry product companies could be helped by this analysis because if they see how consumer laundry habits change over the year, they can learn to market more effectively.

##Kaiyuan Fan

Question: How does weather influence the usage of household in Paris from 2006 to 2010?
For convenient, I will choose two representative date to compare, which is November 1th for winter and June 1th for summer.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
power <- power_data %>%
  separate(Date,into = c("Year", "Month", "Day"), convert = TRUE) %>%
  separate(Time, into = c("Hour", "Minute", "Second"), convert = TRUE) %>%
  mutate(Active_energy_consumed = (Global_active_power*1000/60 - Sub_metering_1 - Sub_metering_2 - Sub_metering_3))

summer1 <- power %>%
  filter(`Day` == 1 & `Month` == 6 & !is.na(Active_energy_consumed)) %>%
  summarize(mean(Active_energy_consumed))
winter1 <- power %>%
  filter(`Day` == 1 & `Month` == 11 & !is.na(Active_energy_consumed)) %>%
  summarize(mean(Active_energy_consumed))
summer2 <- power %>%
  filter(`Day` == 26 & `Month` == 6 & !is.na(Active_energy_consumed)) %>%
  summarize(mean(Active_energy_consumed))
winter2 <- power %>%
  filter(`Day` == 26 & `Month` == 11 & !is.na(Active_energy_consumed)) %>%
  summarize(mean(Active_energy_consumed))
summer3 <- power %>%
  filter(`Day` == 26 & `Month` == 6 & !is.na(Active_energy_consumed)) %>%
  summarize(mean(Active_energy_consumed))
winter3 <- power %>%
  filter(`Day` == 26 & `Month` == 11 & !is.na(Active_energy_consumed)) %>%
  summarize(mean(Active_energy_consumed))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
June1 <- power %>%
  filter(`Month` == 6 | `Month` == 11 & !is.na(Active_energy_consumed))

ggplot(June1, aes(Year, Active_energy_consumed)) +
  geom_point() + 
  facet_wrap(~Month) +
  labs(title = "Comparing Active_energy_consumed in November and June for Different Year")
```

Findings:
Firstly, I compare 3 groups of 2 days. Then I find that for every group, average active energy consumed of a winter day is higher than average active energy consumed of a summer day. 

Then, I make a plot that compare the usage in June and November for different years. I find for every year, November's usage is higher than the June's.
We know paris is cold in winter and warm in summer. In conclusion, weather does influence the usage of household in paris. 

This is an interesting question because it shows how factors outside the dataset can affect measurements that is already in dataset. I use the different month to represent the winter and summer, which is controling the weather. 

I think this analysis would be useful to people who worked in power station.

