---
title: "Lab 7: Team All Stars"
author: "Johnny Tamanaha, Emily Archambault, Eugene Miller, Kaiyuan Fan, Jingzhi Zhang"
date: "10/10/2019"
output: html_document
---

## Individual Sections

## Johnny Tamanaha
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
data <- read_delim("household_power_consumption.txt", delim=";", na = c("","?"), col_types=cols(Date = col_date("%e/%m/%Y"), Time=col_time("%H:%M:%S"), .default=col_double()), locale=locale(decimal_mark="."))
```
```{r}
data_year <- data %>%
  separate(Date, into=c("year", "month", "day"), "-")
avg_global_intensity <- data_year %>%
  group_by(year, Time) %>%
  summarize(avg = mean(Global_intensity, na.rm=TRUE)) %>%
  filter(!is.na(year))
ggplot(avg_global_intensity) +
  geom_point(mapping=aes(x=Time, y=avg, color=year)) +
  labs(title="Average Household Active Power Split by Year") +
  ylab("Active Power (amp)") +
  xlab("Time (Hour:Minutes:Seconds)") +
  scale_color_discrete(name=("Year"))
```
```{r}
avg_global_intensity_stat <- data_year %>%
  group_by(year) %>%
  summarize(avg = mean(Global_intensity, na.rm=TRUE)) %>%
  filter(!is.na(year))
avg_global_intensity_stat
```

Question: How has daily power usage changed from 2006 to 2010?

I first separated the date variable into year, month and day. Then, I grouped the data by year and time and summarized to find the average power usage at each time of each year. Finally, I plotted the time vs the average power and split it by year. To find a summary statistic, I grouped by year and took the average power usage over each year.

My question is interesting because it shows how the general power usage for the people of France changes over time. 

At first I was trying to compare different variables with time between 2006 and 2010. The graphs always looked too cluttered to gain any insight, so I decided to average the variables I chose at each point in time for each year, so that the points would be smoother to look at. After I found the variable that I wanted, I realized that there weren't too many years to plot in the entire data set, so I decided to stick them all on the same plot to show a better time series of the power usage. 

I think that renewable energy companies benefit from my analysis because it shows a slight trend towards less power usage. The exact opposite goes for power companies. 
